---
title: "simulatr on your laptop"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulatr on your laptop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>"
)
```

Here we walk through a simple example of building a `simulatr` specifier object, checking and running the simulation locally, and visualizing the results. We consider estimating the coefficients in a linear regression model via ordinary least squares and lasso, varying the number of samples. 

```{r setup, message = FALSE}
library(simulatr)
library(ggplot2)
```

# 1. Assemble simulation components

- `parameter_grid`. The problem parameters will be the sample size `n`, the dimension `p`, the number of nonzero coefficients `s`, and the value of each nonzero coefficient `beta_val`. In this simulation, we will only vary `n`. We could have therefore put the remainder of the parameters in `fixed_parameters`, but we avoid this to have all problem parameters in `parameter_grid`.
    ```{r}
    parameter_grid <- data.frame(
      n = c(25, 50, 75, 100),      # sample size
      p = 15,                      # dimension
      s = 5,                       # number of nonzero coefficients
      beta_val = 3                 # value of nonzero coefficients
    )
    ```
Next, we want to create the ground truth inferential targets based on the problem parameters. In this case, the inferential target is the coefficient vector. We define a function `get_ground_truth()` that takes as input three of the problem parameters and outputs the coefficient vector:
    ```{r}
    get_ground_truth <- function(p, s, beta_val){
      beta <- numeric(p)
      beta[1:s] <- beta_val
      list(beta = beta)
    }
    ```
We can append these ground truth inferential targets to `parameter_grid` using the helper function `add_ground_truth()`:
    ```{r}
    parameter_grid <- parameter_grid |> add_ground_truth(get_ground_truth)
    ```
Let's take a look at the resulting parameter grid:
    ```{r}
    parameter_grid
    ```
We see the four columns with problem parameters and the fifth with the ground truth coefficient vectors.

- `fixed_parameters`. Here we put only the number of data realizations `B` and the seed to set for the simulation. 
    ```{r}
    fixed_parameters <- list(
      B = 20,                      # number of data realizations
      seed = 4                    # seed to set prior to generating data and running methods
    )
    ```
    
- `generate_data_function`. We define the data-generation function as follows:
    ```{r}
    # define data-generating model based on the Gaussian linear model
    generate_data_f <- function(n, p, ground_truth){
      X <- matrix(rnorm(n*p), n, p, dimnames = list(NULL, paste0("X", 1:p)))
      y <- X %*% ground_truth$beta + rnorm(n)
      data <- list(X = X, y = y)
      data
    }
    ```
Note that we extract the coefficient vector `beta` from the `ground_truth` object. Additionally, we need to call `simulatr_function()` to add a few pieces of information that `simulatr` needs. In particular, we have to give it the argument names of the data-generating function (typically obtained using `formalArgs(generate_data_f)`, as below). We also have to let `simulatr` know whether the data-generating function creates just one realization of the data (`loop = TRUE`) or all `B` at the same time (`loop = FALSE`).
    ```{r}
    # need to call simulatr_function() to give simulatr a few more pieces of info
    generate_data_function <- simulatr_function(
      f = generate_data_f,                        
      arg_names = formalArgs(generate_data_f),    
      loop = TRUE
    )
    ```

- `run_method_functions`. Let's define functions for OLS and lasso:
    ```{r}
    # ordinary least squares
    ols_f <- function(data){
      X <- data$X
      y <- data$y
      lm_fit <- lm(y ~ X - 1)
      beta_hat <- coef(lm_fit)
      results <- list(beta = unname(beta_hat))
      results
    }
    
    # lasso
    lasso_f <- function(data){
      X <- data$X
      y <- data$y
      glmnet_fit <- glmnet::cv.glmnet(x = X, y = y, nfolds = 5, intercept = FALSE)
      beta_hat <- glmnet::coef.glmnet(glmnet_fit, s = "lambda.1se")
      results <- list(beta = beta_hat[-1])
      results
    }
    ```
    Again, we need to call `simulatr_function()` to add a few pieces of information that `simulatr` needs. This time, we need to give it all arguments to the method functions *except the first* (the data itself), which usually will be empty. We also have to let `simulatr` know whether the method functions input just one realization of the data (`loop = TRUE`) or all `B` at the same time (`loop = FALSE`).
    ```{r}
    # create simulatr functions
    ols_spec_f <- simulatr_function(f = ols_f, arg_names = character(0), loop = TRUE)
    lasso_spec_f <- simulatr_function(f = lasso_f, arg_names = character(0), loop = TRUE)
    ```
    Finally, we collate the above method functions into a named list. It is crucial that the list be named. 
    ```{r}
    run_method_functions <- list(ols = ols_spec_f, lasso = lasso_spec_f)
    ```
  
- `evaluation_functions`. Let's evaluate the estimate of `beta` using the RMSE. To accomplish this, we define the following evaluation function:
    ```{r}
    rmse <- function(output, ground_truth) {
      sqrt(sum((output$beta - ground_truth$beta)^2))
    }
    evaluation_functions <- list(rmse = rmse)
    ```
    
# 2. Create a `simulatr` specifier object

This is the easiest step; just pass all four of the above components to the function `simulatr_specifier()`:
```{r}
simulatr_spec <- simulatr_specifier(
  parameter_grid,
  fixed_parameters,
  generate_data_function, 
  run_method_functions,
  evaluation_functions
)
```

# 3. Check and, if necessary, update the `simulatr` specifier object
```{r}
check_results <- check_simulatr_specifier_object(simulatr_spec, B_in = 2)
```
This message tells us that the simulation did not encounter any errors for the first two data realizations. We are free to move on to running the full simulation.
   
# 4. Run the simulation on your laptop

Since this example simulation is small, we can run it on your laptop in RStudio:
```{r}
sim_results <- check_simulatr_specifier_object(simulatr_spec)
```

# 5. Summarize and/or visualize the results

Let's take a look at the results:
```{r}
sim_results$metrics
```
We have both the mean and Monte Carlo standard error for the metric (RMSE) for each method in each problem setting. We can plot these as follows:

```{r, fig.align='center', fig.width=6, fig.height = 4}
sim_results$metrics |>
  ggplot(aes(x = n, 
             y = mean, 
             ymin = mean - 2*se, 
             ymax = mean + 2*se, 
             color = method)) +
  geom_point() + 
  geom_line() + 
  geom_errorbar(width = 1) +
  labs(x = "Sample size",
       y = "RMSE") + 
  theme(legend.position = "bottom")
```

It looks like lasso performs better for small sample sizes but OLS performs better for large sample sizes.